---
title: "Homework 5"
author: "Ekaterina Hofrenning"
date: '2023-11-02'
output: github_document
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(purrr)
library(broom)
```

# Problem 2
This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm. Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

- Start with a dataframe containing all file names; the list.files function will help
- Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe
- Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary
- Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r, warning=FALSE, message=FALSE}
sets <- list.files("data", full.names = T)

files <- purrr::map(sets, read_csv) %>% 
  bind_rows() %>%
  mutate(control_arm = as.factor(if_else(row_number() %in% c(1:10), "Control", "Experimental")),
         ID = row_number()) %>%
  relocate(control_arm, ID, .before = "week_1") %>%
  pivot_longer(cols = week_1:week_8, names_to = "week", values_to = "value") %>%
  separate(week, c("text", "week")) %>%
  mutate(week = as.integer(week)) %>%
  select(-text)
```

```{r, warning=FALSE, message=FALSE}
ggplot(files, aes(x = week, y = value, group = ID, color = control_arm)) +
  geom_line() +
  labs(x = "Week", y = "Value", color = "Arm")
```

As seen in the plot above, the experimental arm has a higher average value than the control arm. Additionally, the experimental arm's values rise as the weeks go on, whereas the control arms appear steady on average.


# Problem 3

When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

First set the following design elements:

- Fix n=30
- Fix σ=5
- Set μ=0. 

Generate 5000 datasets from the model using x∼Normal[μ,σ].

For each dataset, save μ̂ and the p-value arising from a test of H:μ=0 using α=0.05.

Repeat the above for μ={1,2,3,4,5,6}, and complete the following:

- Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. Describe the association between effect size and power.
- Make a plot showing the average estimate of μ̂ on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of μ̂ only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. Is the sample average of μ̂ across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?


```{r, warning=FALSE, message=FALSE}
# Define t-test function w/ set parameter
t_test <- function(mu = 0) {
  sample <- tibble(rnorm(n = 30, mean = mu, sd = 5))
  
  results <- t.test(sample) %>% 
    broom::tidy() %>% 
    select(estimate, p.value)
  
  results
}

# mean = 0
mean_0 <- expand_grid(mean = 0, 
                      iteration = 1:5000) %>% 
  mutate(results = map(mean, t_test)) %>% 
  unnest(results)

# mean = 1,2,3,4,5,6
mean_all <- expand_grid(mean = 1:6, 
                        iteration = 1:5000) %>% 
  mutate(results = map(mean, t_test)) %>% 
  unnest(results)
```


```{r, warning=FALSE, message=FALSE}
mean_all %>%
  group_by(mean) %>% 
  summarize(prop_rej = sum(p.value < 0.05)/5000) %>% 
  ggplot(aes(x = mean, y = prop_rej)) +
  scale_x_continuous(limits = c(1,6), breaks = seq(1,6,1)) + 
  geom_point() + 
  geom_path() +
  labs(x = "True Mean", y = "Power (Proportion Null was Rejected)")
```

When the true mean increases, the power increases. The rate of power increases tapers off as it approaches 1.

```{r, warning=FALSE, message=FALSE}
rej_est <- mean_all %>% 
  filter(p.value < 0.05) %>% 
  group_by(mean) %>% 
  summarize(ave_est = mean(estimate, na.rm = T)) %>% 
  ungroup()

full_est <- mean_all %>% 
  group_by(mean) %>% 
  summarize(ave_est = mean(estimate, na.rm = T)) %>% 
  ungroup()
  
ggplot(full_est, aes(x = mean, y = ave_est)) +
  geom_line(data = full_est, aes(color = "red")) +
  geom_line(data = rej_est, aes(color = "blue")) +
  scale_color_manual(name = " ", values = c("red" = "red", "blue" = "blue"),
                     labels = c('Rejected Estimates','All Estimates')) +
  geom_point(data = full_est, color = "red") +
  geom_point(data = rej_est, color = "blue") +
  scale_x_continuous(limits = c(1,6), breaks = seq(1,6,1)) +
  labs(x = "True Mean", y = "Average Estimated Mean", title = "All and Rejected Mean Estimates")
```

For all estimates, the true mean is approximately equal to the average estimated mean. But, for rejected estimates, this is only the case for means larger than 4-- for those less than 4, the average estimated mean is larger than the true mean. This is because the power of the test is smaller when the true mean is smaller. 

